Block bootstrap consistency\\under weak assumptions

### 2016-09-14 Paper.tex
[#27] Tweak centering term in Brownian Motion
Related to issue #20, I think it makes sense to center on a different
variable. This commit changes the centering term and revises my letter
to the referee.
[#27] Make small changes to wording in main paper

\newpage
\noindent Block bootstraps, e.g.\ the Moving Blocks
(\citealp{Kun:89}, and \citealp{LiS:92}), Circular Block \citep{PoR:92}, and Stationary
Bootstraps \citep{PoR:94}, have become popular in Economics, partly
because they do not require the researcher to make parametric
assumptions about the data generating process.  They are valid under
general weak dependence and moment conditions.  Some recent papers
(\citealp{GoW:02}, and \citealp{GoJ:03}) relax the dependence and moment
conditions of the original papers to fit with the Near-Epoch-Dependence
(\ned) assumptions commonly used in econometrics.%
\footnote{\citet{GoW:02} show that these bootstrap methods can be
  applied to heterogeneous $L_{2+\delta}$-\ned\ processes of size
  $-2(r-1)/(r-2)$ on a strong mixing sequence of size
  $-r(2+\delta)/(r-2)$, where $r > 2$ and $\delta >0$, when the
  original series has uniformly bounded $3r$-moments.  \citet{GoJ:03}
  relax these conditions to $L_{2+\delta}$-\ned\ of size $-1$ and
  $r+\delta$ moments for the original series, and size
  $-(2+\delta)(r+\delta)/(r-2)$ for the underlying mixing series.
  Both papers require that the expected block length grow with $n$ and
  be $o(n^{1/2})$.  \cite{GoP:11} discuss these issues further.}%
\footnote{An array $\{X_{nt}\}$ is an $L_{\rho}$-\ned\ process on a
  mixing array $\{V_{nt}\}$ if
  \begin{equation}
    \| X_{nt} - \E(X_{nt}
    \mid V_{n,t-m},\dots,V_{n,t+m}) \|_{\rho} \leq d_{nt} v_m
  \end{equation}
  with $v_m \to 0$ as $m \to \infty$ and $\{d_{nt}\}$ an array of
  positive constants.  It is of size $-\gamma$ if $v_m = O(m^{-\gamma
    - \delta})$ for all $\delta>0$.  Dropping the index ``$n$'' gives
  the series definition.  Note that the underlying strong and uniform mixing arrays
  are not required to be stationary.} %
But these conditions are still stronger than
required for a \clt\ to hold; \citet{Jon:97} has established the \clt\
under $L_2$-\ned\ with smaller size and moment
restrictions.%
\footnote{\citet{Jon:97} proves that the \clt\ holds for averages of
  $L_2$-\ned\ processes of size $-1/2$ on a strong mixing series of
  size $-r/(r-2)$, $r > 2$ and the original series having bounded
  $r$-moments.} %
This paper shows that these block bootstrap
methods consistently estimate the distribution of the sample mean
under \citepos{Jon:97} assumptions and show that an \fclt\ holds as
well.%
\footnote{\citet{Rad:96} proves consistency for the Moving Blocks
  Bootstrap for any stationary strong mixing sequence that satisfies
  the \clt. This paper uses a similar method of proof to his, but also
  accommodates nonstationary sequences and the Stationary
  Bootstrap.} %
It also relaxes \citepos{GoW:02} and \citepos{GoJ:03} requirement that
the expected block length be $o(n^{1/2})$ to the original papers'
requirement that it be $o(n)$, although their stronger assumption is
necessary for the bootstrap to be consistent under some forms of
heterogeneity.

@@

The assumption on the dispersion of the individual means,
$(\mu_{nt} - \bar\mu_n)^2$, is the same as \citepos{GoW:02} and
\citepos{GoJ:03} under their additional requirement that the
expected block length be $o(\sqrt{n})$. When
the blocks are larger than $o(\sqrt{n})$, a case that \citet{GoW:02}
and \citet{GoJ:03} do not consider, the dispersion must be smaller.
To see these implications of this assumption, it is helpful to
consider a simple example: a stochastic process with a single break of
size $\Delta_n$ at time $T_n$, implying
\begin{equation*}
\mu_{nt} =
\begin{cases}
  \mu & \text{if } t \leq T_n \\
  \mu + \Delta_n & \text{otherwise}
\end{cases}
\end{equation*}
A small amount of algebra gives the relationship,
\begin{equation*}
  \sum_{t=1}^n (\mu_{nt} - \bar \mu_n)^2
  = T_n (n - T_n) \Delta_n^2 / n \equiv D_n;
\end{equation*}
we have defined $D_n$ to minimize notation in the subsequent
discussion.  For the Moving and Circular Block Bootstraps,
Theorem~\ref{T1}'s Assumption 2 holds if $D_n = o(n / M_n^2)$, and for
the Stationary Bootstrap the same relationship must hold for the
expected block length. If $\Delta_n = \Delta_0$ and $T_n = \pi n$ for
some constants $\Delta_0$ and $\pi$ and for all $n$, then
$D_n \approx \Delta_0^2 \, c (1 - c) \, n$ and Assumption~2 fails to
hold. These values of $\Delta_n$ and $T_n$ describe a scenario where
there is a break large enough that its timing could be precisely
estimated \citep[as in][]{Bai:94}, but the break is ignored and the
bootstrap is applied to the entire dataset.

We can also use the same DGP to produce scenarios where
Theorem~\ref{T1}'s Assumption~2 is satisfied.  The first mimics a
setting where the break is estimated and the bootstrap is applied to
the post-break sample, but a small number of pre-break observations
are accidentally included in the data set as well. Here
$\Delta_n = \Delta_0$ and $T_n = T_0$ for some finite $T_0$. In this
case, Assumption 2 holds hold as long as $M_n = o(\sqrt{n})$.
Similarly, if the break is moderate with unknown timing in the sense
of \citet{ElM:07}, so $\Delta_n = \Delta / \sqrt{n}$ and
$T_n = \pi n$, then $D_n \approx \Delta c (1 - c)$ and Assumption 2
again holds as long as $M_n = o(\sqrt{n})$. Consequently, using block sizes
of length $o(\sqrt{n})$ ensures that Theorem 1's assumptions on the
dispersion of the $\mu_{nt}$ hold when there is a moderate amount of
heterogeneity---either because the sample is chosen to lie after a
large break estimated break or because any breaks are smaller in
magnitude and, consequently, potentially difficult to detect. When the
degree of heterogeneity is very small or nonexistent, so
$\sum_t (\mu_{nt} - \bar \mu_n)^2$ grows even slower, then larger
block lengths can also satisfy this condition.

@@

Theorem~\ref{T1} can also be extended to give an \fclt\
using arguments from \citet{JoD:00b}. We show in
Theorem~\ref{T2} that the partial sum
of the bootstrapped process obeys an \fclt\ and can be used to derive
critical values for other test statistics under the same assumptions
as Theorem~\ref{T1}. For this result, define the
following partial sums,
\begin{align*}
W_n(\gamma) &=
\tfrac{1}{\sqrt{n}} \sum_{t=1}^{\lfloor \gamma n \rfloor}
(X_{nt} - \bar\mu_n) &&\text{and}&
W_n^{*}(\gamma) &= \tfrac{1}{\sqrt{n}}
\sum_{t=1}^{\lfloor \gamma n \rfloor}
(X^*_{nt} - \bar\mu_n^{*}).
\end{align*}
Also let $W$ denote standard Brownian Motion and $\sigma W$ denote
Brownian Motion scaled by the constant $\sigma$.

>>>>

\newpage
\noindent Block bootstraps, e.g.\ the Moving Blocks
(\citealp{Kun:89}, and \citealp{LiS:92}), Circular Block \citep{PoR:92}, and Stationary
Bootstraps \citep{PoR:94}, have become popular in Economics, partly
because they do not require the researcher to make parametric
assumptions about the data generating process.  They are valid under
general weak dependence and moment conditions.  Some recent papers
(\citealp{GoW:02}, and \citealp{GoJ:03}) relax the dependence and moment
conditions of the original papers to fit with the Near-Epoch-Dependence
(\ned) assumptions commonly used in econometrics.%
\footnote{\citet{GoW:02} show that these bootstrap methods can be
  applied to heterogeneous $L_{2+\delta}$-\ned\ processes of size
  $-2(r-1)/(r-2)$ on a strong mixing sequence of size
  $-r(2+\delta)/(r-2)$, where $r > 2$ and $\delta >0$, when the
  original series has uniformly bounded $3r$-moments.  \citet{GoJ:03}
  relax these conditions to $L_{2+\delta}$-\ned\ of size $-1$ and
  $r+\delta$ moments for the original series, and size
  $-(2+\delta)(r+\delta)/(r-2)$ for the underlying mixing series.
  Both papers require that the expected block length grow with $n$ and
  be $o(n^{1/2})$.  \cite{GoP:11} discuss these issues further.}%
\footnote{An array $\{X_{nt}\}$ is an $L_{\rho}$-\ned\ process on a
  mixing array $\{V_{nt}\}$ if
  \begin{equation}
    \| X_{nt} - \E(X_{nt}
    \mid V_{n,t-m},\dots,V_{n,t+m}) \|_{\rho} \leq d_{nt} v_m
  \end{equation}
  with $v_m \to 0$ as $m \to \infty$ and $\{d_{nt}\}$ an array of
  positive constants.  It is of size $-\gamma$ if $v_m = O(m^{-\gamma
    - \delta})$ for all $\delta>0$.  Dropping the index ``$n$'' gives
  the series definition.  Note that the underlying strong and uniform mixing arrays
  are not required to be stationary.} %
But these conditions are still stronger than
required for a \clt\ to hold; \citet{Jon:97} has established the \clt\
under $L_2$-\ned\ with smaller size and moment
restrictions.%
\footnote{\citet{Jon:97} proves that the \clt\ holds for averages of
  $L_2$-\ned\ processes of size $-1/2$ on a strong mixing series of
  size $-r/(r-2)$, $r > 2$ and the original series having bounded
  $r$-moments.} %
This paper shows that these block bootstrap
methods consistently estimate the distribution of the sample mean
under \citepos{Jon:97} assumptions and show that an \fclt\ holds as
well.%
\footnote{\citet{Rad:96} proves consistency for the Moving Blocks
  Bootstrap for any stationary strong mixing sequence that satisfies
  the \clt. This paper uses a similar method of proof to his, but also
  accommodates nonstationary sequences and the Stationary
  Bootstrap.} %
It also relaxes \citepos{GoW:02} and \citepos{GoJ:03} requirement that
the expected block length be $o(n^{1/2})$ to the original papers'
requirement that it be $o(n)$, although, as we will show, their
stronger assumption is necessary for the bootstrap to be consistent
under some forms of heterogeneity.

@@

The assumption on the dispersion of the individual means,
$(\mu_{nt} - \bar\mu_n)^2$, is the same as \citepos{GoW:02} and
\citepos{GoJ:03} under their additional requirement that the
expected block length be $o(\sqrt{n})$. When
the blocks are larger than $o(\sqrt{n})$, a case that \citet{GoW:02}
and \citet{GoJ:03} do not consider, the dispersion must be smaller
than they require.
To see these implications of this assumption, it is helpful to
consider a simple example: a stochastic process with a single break of
size $\Delta_n$ at time $T_n$:
\begin{equation*}
\mu_{nt} =
\begin{cases}
  \mu & \text{if } t \leq T_n \\
  \mu + \Delta_n & \text{otherwise}.
\end{cases}
\end{equation*}
A small amount of algebra gives the relationship,
\begin{equation*}
  \sum_{t=1}^n (\mu_{nt} - \bar \mu_n)^2
  = T_n (n - T_n) \Delta_n^2 / n \equiv D_n;
\end{equation*}
we have defined $D_n$ to minimize notation in the subsequent
discussion.  For the Moving and Circular Block Bootstraps,
Theorem~\ref{T1}'s Assumption 2 holds if $D_n = o(n / M_n^2)$, and for
the same relationship must hold for the expected block length for the
Stationary Bootstrap. If $\Delta_n = \Delta_0$ and $T_n = \pi n$ for
some constants $\Delta_0$ and $\pi$ and for all $n$, then
$D_n \approx \Delta_0^2 \, c (1 - c) \, n$ and Assumption~2 fails to
hold. These values of $\Delta_n$ and $T_n$ describe a scenario where
there is a break large enough that its timing could be precisely
estimated \citep[as in][]{Bai:94}, but the break is ignored and the
bootstrap is applied to the entire dataset.

To see when Theorem~\ref{T1}'s Assumption~2 is satisfied, consider two
other parametrizations of this DGP. First, let $\Delta_n = \Delta_0$
and $T_n = T_0$ for some finite $T_0$. This mimics a setting where a
researcher intends to apply the bootstrap to a post-break sample, but
mistakenly includes a small number of pre-break observations as
well. In this case, Assumption 2 holds hold as long as
$M_n = o(\sqrt{n})$.  Similarly, suppose
$\Delta_n = \Delta / \sqrt{n}$ and $T_n = \pi n$, which mimics a
setting with a moderate break with unknown timing.  Then
$D_n \approx \Delta c (1 - c)$ and Assumption 2 again holds as long as
$M_n = o(\sqrt{n})$. Consequently, using block sizes of length
$o(\sqrt{n})$ ensures that Theorem 1's assumptions on the dispersion
of the $\mu_{nt}$ hold when there is a moderate amount of
heterogeneity---either because the sample is chosen to lie after a
large break estimated break or because any breaks are smaller in
magnitude and, consequently, potentially difficult to detect. When the
degree of heterogeneity is very small or nonexistent, so
$\sum_t (\mu_{nt} - \bar \mu_n)^2$ grows even slower, then larger
block lengths can also satisfy this condition.

@@

Theorem~\ref{T1} can also be extended to give an \fclt\
using arguments from \citet{JoD:00b}. We show in
Theorem~\ref{T2} that the partial sum
of the bootstrapped process obeys an \fclt\ and can be used to derive
critical values for other test statistics under the same assumptions
as Theorem~\ref{T1}. For this result, define the
following partial sums,
\begin{align*}
W_n(\gamma) &=
\tfrac{1}{\sqrt{n}} \sum_{t=1}^{\lfloor \gamma n \rfloor}
(X_{nt} - \bar\mu_n) &&\text{and}&
W_n^{*}(\gamma) &= \tfrac{1}{\sqrt{n}}
\sum_{t=1}^{\lfloor \gamma n \rfloor}
(X^*_{nt} - \E^* \bar X_n^*).
\end{align*}
Also let $W$ denote standard Brownian Motion and $\sigma W$ denote
Brownian Motion scaled by the constant $\sigma$.
